A weakly structured stem for human origins in Africa

### 2022-03-11 paper.tex
Made a few small edits to the introduction

Archaeological sites from the Middle Stone Age (approx. 300ka-40ka) are widely
distributed across Africa, and are particularly well represented in the
northern, eastern and southern parts of the continent. Similarly, fossil crania
such as those from the sites of Jebel Irhoud \citep{Hublin2017-cq}, Herto
\citep{White2003-bk} and Klasies River \citep{Deacon1995-rx} demonstrate that
anatomically derived \emph{Homo sapiens} features were also present across the
continent during this period. It has been difficult to reconcile these lines of
evidence with evidence from genomics, which have suggested a predominantly
tree-like model of recent population divergence from a single ancestral
population. It is unclear whether fossil specimens and archaeological sites
represent populations which contributed to our ancestors as population
precedents, or were local ``dead-ends'' from which contemporary \emph{Homo
sapiens} do not descend. 
Recently,
attempts to reconcile genetic and
paleoanthropological data include proposals for a Pan-African origin of
\emph{Homo sapiens} by which populations in many regions of the continent
contributed to the formation of Homo sapiens beginning at least 300ka
\citep{Stringer2016-mj,Scerri2018-nl,Scerri2019-xg}.

Genetic models have been hampered in their contribution to this discussion
because they primarily assume (or, at least, have been tested under) a
tree-like model of isolation-with-migration. Alternative theoretical scenarios
have been proposed, such as stepping stone models \citep{Arredondo2021-qa} or
population coalescence and fragmentation \citep{Scerri2019-xg}. These
approaches are more challenging to interpret and fit to data.
However, new population genetic tools now allow for inference involving tens to
hundreds of genomes from multiple populations and greater complexity
\citep{Kamm2020-vn,Ragsdale2019-nt,Speidel2019-nj}. Inspired by evidence for
Neanderthal admixture with humans in Eurasia, several recent articles
have shown that introducing an archaic hominin ghost population contributing to African
populations in the period surrounding the Out-of-Africa migration event
substantially improves the description of genetic data relative to
single-origin models
\citep{Plagnol2006-lt,Hammer2011-bx,Hsieh2016-gk,Hey2018-pw,Ragsdale2019-nt,Durvasula2020-td,Lorente-Galdos2019-vz,Durvasula2020-td}.
This has driven speculation about the geographic range of this ghost
population, possible links to specific fossils, and the possibility of
finding ancient DNA evidence \citep{Hsieh2016-gk}). However, these prior
articles share two weaknesses. First, they only contrast a single-origin model
with an archaic hominin admixture model, leaving out other plausible models
\citep{Henn2018-rf} (Figure~\ref{fig:proposed-models}).
Second, they
focus on a small subset of African diversity, either because of small sample
sizes (2-5 genomes) or because they rely on 1000 Genomes data which only
recruited populations of recent West African or Bantu-speaking ancestry
(Figure~\ref{fig:diversity}C). 
While ancient DNA from Eurasia has helped us understand early human history
outside of Africa, there is no comparably ancient DNA to elucidate early history
in Africa \citep{Lipson2022-xf}.

>>>>

Archaeological sites from the Middle Stone Age (approx. 300-40ka) are widely
distributed across Africa, and are particularly well represented in the
northern, eastern and southern parts of the continent. Similarly, fossils 
such as those from the sites of Jebel Irhoud, Morocco \citep{Hublin2017-cq}, Herto, Ethiopia
\citep{White2003-bk} and Klasies River, South Africa \citep{Deacon1995-rx} demonstrate that
anatomically derived \emph{Homo sapiens} features were also present across the
continent during this period. It has been difficult to reconcile these lines of
evidence with evidence from genomics, which have suggested a predominantly
tree-like model of recent population divergence from a single ancestral
population. It is unclear whether fossil specimens and archaeological sites
represent populations which contributed to our ancestors as population
precedents, or were local ``dead-ends'' from which contemporary \emph{Homo
sapiens} do not descend. 
Recently,
attempts to reconcile genetic and
paleoanthropological data include proposals for a Pan-African origin of
\emph{Homo sapiens} by which populations in many regions of the continent
contributed to the formation of /emph{Homo sapiens} beginning at least 300ka
\citep{Stringer2016-mj,Scerri2018-nl,Scerri2019-xg}.

Genetic models have been hampered in their contribution to this discussion
because they primarily assume (or, at least, have been tested under) a
tree-like model of isolation-with-migration. Alternative theoretical scenarios
have been proposed, such as stepping stone models \citep{Arredondo2021-qa} or
population coalescence and fragmentation \citep{Scerri2019-xg}, but these
approaches are more challenging to interpret and fit to data.
However, new population genetic tools now allow for inference involving tens to
hundreds of genomes from multiple populations and greater complexity
\citep{Kamm2020-vn,Ragsdale2019-nt,Speidel2019-nj}. Inspired by evidence for
Neanderthal admixture with humans in Eurasia, several recent articles
have shown that introducing an archaic hominin ghost population contributing to African
populations in the period surrounding the Out-of-Africa migration event
substantially improves the description of genetic data relative to
single-origin models
\citep{Plagnol2006-lt,Hammer2011-bx,Hsieh2016-gk,Hey2018-pw,Ragsdale2019-nt,Durvasula2020-td,Lorente-Galdos2019-vz,Durvasula2020-td}.
This has driven speculation about the geographic range of this ghost
population, possible links to specific fossils, and the possibility of
finding ancient DNA evidence \citep{Hsieh2016-gk}. However, these prior
articles share two weaknesses. First, they only contrast a single-origin model
with an archaic hominin admixture model, leaving out other plausible models
\citep{Henn2018-rf} (Figure~\ref{fig:proposed-models}).
Second, they
focus on a small subset of African diversity, either because of small sample
sizes (2-5 genomes) or because they rely on 1000 Genomes data which only
recruited populations of recent West African or Bantu-speaking ancestry
(Figure~\ref{fig:diversity}C). 
While ancient DNA from Eurasia has helped us understand early human history
outside of Africa, there is no comparably ancient DNA to elucidate early history
in Africa \citep{Lipson2022-xf}.
